### 2016-09-29 outline.rmd
hansen

# Introduction

@@

Something similar applies to recommendation systems. The system can remember very many things.
It may know everything you bought in the last few years on that site, and it can easily see what you would normally buy and add those as suggestions.
But a recommendation system finds it much harder to suggest items that you have not forgotten but may have never bought.

If we are interested in lists from various angles -- asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or even in terms of different cultural practices of listing [@Jullien_TBA] -- then contemporary lists practices such as recommendations figure as sites of intersection between different forms.

Lists also raise the question of how we institute our approach to them.
While lists such as recommendations intersect strongly with meaning and sense-making practices, they are also somewhere refractory to methods of humanities and social science research.
For on the one hand, lists are highly associative. They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that limit semiotic readings.
For on the other hand, as operational entities, they can be treated ethnomethodologically, as the production of ('the order of social settings through ...  shared sense making practices.')[https://en.wikipedia.org/wiki/Ethnomethodology]  The question for me here is whether contemporary lists can still be understood as produced through 'shared sense-making practices.' 
All lists are forms of order, since even random ordering is operationally important, but what counts as order might historically transformable, in which case the order of social settings or the problem of social order might be deeply historically contingent. 
The problem of order lies at the core of lists making, and our relation to lists can be seen as a test of the way that order operates more generally. 
Much of my paper, and my responses to the presentation I saw from a data scientist  will focus on how the ordering that is done in a recommendation lists relates to the problem of social order more generally. 
- The main narrative of the presentation concerned a shift from a set-based model for making lists to a predictive, probabilistic way of making lists in almost realtime and dealing with lists on a scale that is often difficult to grasp concretely.
I respond to the presentation in two ways -- firstly by recounting, contextualising and commenting on some of the main steps as a example of the sense-making done by data scientists and customer analytics teams in many places; and secondly, by trying to step into the list-making practices  as an occasion to intervene and participate in the list-making practices they engage in.

>>>>

# Introduction: the presentation of a presentation

@@

Something similar applies to recommendation systems. The system can remember very many things.
It may know everything you bought in the last few years on that site, and it can easily see what you would normally buy and add those as suggestions.
But a recommendation system finds it much harder to suggest items that you have not forgotten but may have never bought. If you do buy recommended items, then the recommender system has been effective. It has led to a 'conversion.'

If we are interested in lists from various angles -- as asignifying semiotics as Maurizio Lazzarato calls it in his _Signs and Machines_ [@Lazzarato_2014] or as elements in 'a new order of spatio-temporal continuity for forms of economic, political and cultural life' as Celia Lury terms in her account of topological turn [@TBA], or even in terms of different cultural practices of listing [@Jullien_TBA] -- then contemporary list practices such as recommendations figure as sites of intersection between different forms.

Lists also raise the question of how we institute our approach to them.
While lists such as recommendations intersect strongly with other meaning and sense-making practices in everyday life and popular culture, they are also somewhere refractory to methods of humanities and social science research.
For on the one hand, lists are highly associative. They can be interpreted or decoded semiotically, but they also exhibit great variations in textuality that limit semiotic readings. For on the other hand, as operational entities, they can be treated ethnomethodologically, as the production of ('the order of social settings through ...  shared sense making practices.')[https://en.wikipedia.org/wiki/Ethnomethodology]

The question for me here is whether contemporary lists can still be understood as produced through 'shared sense-making practices.' All lists are forms of order, since even random ordering is operationally important, but what counts as order might historically transformable, in which case the order of social settings or the problem of social order might be deeply historically contingent. The problem of order lies at the core of lists making, and our relation to lists can be seen as a test of the way that order operates more generally.

Much of my paper, and my responses to the presentation I saw from a data scientist  will focus on how the ordering that is done in a recommendation lists relates to the problem of social order more generally.  The main narrative of the presentation concerned a shift from a set-based model for making lists to a predictive, probabilistic way of making lists in almost realtime and dealing with lists on a scale that is often difficult to grasp concretely. The change in list-making associated with learned models is increasingly widespread, and
I respond to the presentation in two ways -- firstly by recounting, contextualising and commenting on some of the main steps as a example of the sense-making done by data scientists and customer analytics teams in many places; and secondly, by trying to step into the list-making practices  as an occasion to intervene and participate in the list-making practices they engage in. This is no mere methodological problem if, as Mark Hansen has recently suggested,

> predictive analytics are discoveries of micrological propensities that are not directly correlated with human understanding and affectivity and that do not by themselves cohere into clearly identifiable events [@Hansen_2015b, 111-112]

The question exercising much contemporary work on data, on media, platforms and finance pivots on the 'micrological propensities' that might be re-ordering lists. Hansen's argument here is that these propensities make probabilities into something real, some actually in things:

>Whatever explanatory and causal value predictive analytics of large datasets have is, I suggest, ultimately rooted in this ontological transformation whereby probabilities are understood to be expressions of the actual propensity of things [@Hansen_2015b, 120]

As the quote shows, he presents this as an 'ontological transformation.' The practical and indeed experimental question for me here is whether such transformations can be detected and articulated in the prosaic setting of shopping lists and  shopping baskets.
